Using tours to visually investigate properties of new projection pursuit indexes with application to problems in physics

### 2018-12-13 scagnostics.Rmd
Added non-smoothed for comparison in smoothing plot, short statement about scaling of computing time.

This is particularly interesting for the scagnostics indexes skinny and stringy which we found to be most noisy among the indexes considered. Figure \ref{fig:spiralScagSmoothing} studies the potential of these two smoothing approaches, using a the tour path between noise variables of the spiral1000 dataset using different $\epsilon$ and $\beta$ values. Both methods appear to be promising in smoothing the function. Because the scagnistics are fast to compute, wither of these methods is feasible.

@@

```{r spiralScagSmoothing, fig.cap="Comparing the traces of the scagnostics indices skinny and stringy when smoothing the index values, either by averaging over the index value after jittering the projection by some angle $\\epsilon$ (full line) or after jittering the projected datapoints with some amount $\\beta$ (dashed line). While the behaviour can be somewhat improved, sharp jumps can be observed in all settings considered."}

@@

scagSmoothMelt <- scagSmoothMelt %>%
  filter(variable != "convex") %>%
  mutate(method = factor(method, levels = c("jitterAngle", "jitterPoints")))
ggplot(scagSmoothMelt, aes(x=t, y=value)) +
  geom_line(aes(linetype=factor(method))) + 
  theme(legend.position="none") +
  ylim(0,1) +
  facet_grid(alpha~variable) +
  xlab("Sequence of projections (t)") +
  ylab("PPI value")

>>>>

This is particularly interesting for the scagnostics indexes skinny and stringy which we found to be most noisy among the indexes considered. Figure \ref{fig:spiralScagSmoothing} studies the potential of these two smoothing approaches, using a the tour path between noise variables of the spiral1000 dataset using different $\epsilon$ and $\beta$ values. Both methods appear to be promising in smoothing the function. Because the scagnistics are fast to compute, either of these methods is feasible. For this example we have smoothed over 10 randomly selected jittered views, computing time increases linearly with the number of randomly jittered views, as this is mostly determined by the time needed to evaluate the scagnostics indexes which is done separately for each view.

@@

```{r spiralScagSmoothing, fig.cap="Comparing the traces of the scagnostics indices skinny and stringy when smoothing the index values, either by averaging over the index value after jittering the projection by some angle $\\epsilon$ (full line) or after jittering the projected datapoints with some amount $\\beta$ (dashed line). While the behaviour can be somewhat improved, sharp jumps can be observed in all settings considered. For comparison the red line in the background shows the trace without any smoothing applied."}

@@

#getting non-smoothed result
load("cache/V1V2toV3V4.rda")
noSmoothing <- fullRes %>%
  filter(size==1000) %>%
  filter(name=="spiral") %>%
  select(skinny, stringy, t) %>%
  gather(variable, value, -t)

scagSmoothMelt <- scagSmoothMelt %>%
  filter(variable != "convex") %>%
  mutate(method = factor(method, levels = c("jitterAngle", "jitterPoints")))
ggplot(scagSmoothMelt, aes(x=t, y=value)) +
  geom_line(data=noSmoothing, color="red", alpha=0.5) +
  geom_line(aes(linetype=factor(method))) + 
  theme(legend.position="none") +
  ylim(0,1) +
  facet_grid(alpha~variable) +
  xlab("Sequence of projections (t)") +
  ylab("PPI value")
